---
title: "Assignment2 Exercise 2"
output: pdf_document
---

Exercise 2

First, we load data from "search.txt"
```{r setup, include=FALSE}
search = read.table("search.txt",header = TRUE);
attach(search)
```
Task a

To show the randomization of the data, we use N to present the dependent variable, use I to present the treatment, and use B to present the block.
```{r}
N = 1; I = 3;B = 5;
for (i in 1:B){
  print(sample(1:(N*I)))
}
```
The rows show the skill level of each student.

Task 2

Below is the graphic summaries of the data
```{r}
interaction.plot(skill,interface,time)
```

```{r}
interaction.plot(interface,skill,time)
```
The lines are not parallel, which means there is interaction between interface and skill.

Task 3

We use a 2-way anova test to setimate the hypothesis. 
```{r}
search$skill = as.factor(search$skill)
search$interface = as.factor(search$interface)

searchaov = lm(time~interface + skill,data=search)
anova(searchaov)
```
The P-value is 0.013, which means we can reject the null hypotheis that the mean of interfaces are the same.

```{r}
searchaov2 = lm(time~interface + skill, data = search)
summary(searchaov2)
```
The time it takes a level 3 user to find a product if the website uses interface 2 is:
    15.01+3.03+2.70 = 20.74
    
Task 4

```{r}
par(mfrow=c(1,2));par(pin=c(2,2))
qqnorm(residuals(searchaov2))
qqline(residuals(searchaov2))
plot(fitted(searchaov2),residuals(searchaov2))
```

The qq-plot seems a bit deviated in the extremes but it could be normal and the residuals don't change systematically with the fitted values. So we can assume that the populations have equal variances. 

Task 5

To test whether there is an effect of interface, we use the Friedman test.
```{r}
friedman.test(search$time,search$interface,search$skill)
```
P-value is 0.04, so we reject the null hypothesis, which we can say the interface makes an effect

Task 6
```{r}
search$interface = as.factor(search$interface)
searchaov = lm(time~interface,data=search)
anova(searchaov)
```
The P-value is 0.096 and we can not reject that the means are the same for the different interfaces.
Then we use one-way-anova assumes that the data come from normal population and the variances are equal.
```{r}
qqnorm(residuals(searchaov))
qqline(residuals(searchaov))
```
The population is not normal and the assumption is not met.
